\bf Nonparametric Bootstrap Kolmogorov-Smirnov Test with Stationary Time Series

### 2023-11-30 manuscript/manuscript.tex
11/29 Notes

The Kolmogorov-Smirnov (KS) statistic is widely used to test if a sample is
from a given distribution.

@@

\paragraph{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ from a certain distribution $F(\theta)$,
and use the block bootstrap KS test to test if $X_t$ is generated from 
distribution $F$ with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformally distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
The results of the method depend on the sample size, so we must find the minimum
sample size required for the block bootstrap KS test to work.
We can implement the procedure described in methods with the true distribution
$F ~ N$, hypothesize distribution $F_0 ~ N$, and true $\mu = 8$ and 
$\sigma = 8$.
We replicate this procedure 1000 times to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.
Then, we repeat the same procedure with $F ~ \Gamma$.

We constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, this 
indicates that the time series follows the null distribution. 
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$. We can also
compute the rate of rejection $\#P < \alpha$ for different values of
$\alpha$, as well as a confidence interval
for this proportion.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/sim_800_normal_-0.4_0}
  \caption{Q-Q p}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/zoom_800_normal_-0.4_0}
  \caption{}
  \label{fig:mu}
\end{figure}

\paragraph{Power}
We must also evaluate if this method works when $H_0$ is false. We can use
the block bootstrap KS test to test if $X_t$ is generated from some other 
distribution $G$ with some unknown $\theta$. In this scenario, if the test is 
indeed powerful,
we would ideally want $\beta$, or the probability of rejection to be 1, but we
expect it to be generally high. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-normal, and the rate
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

If we can show that the test correctly fails to reject $H_0$ under the null
distribution, and correctly rejects $H_0$ for an alternate distribution, we can
state tht the method works.

@@

\subsection{Precipitation in Mansfield Connecticut}
Precipitation annual maximum in inches series data was retrieved from the 
National Oceanic and Atmospheric Administration.
%\mc{https://hdsc.nws.noaa.gov/pfds/pfds_map_cont.html?bkmrk=ct}

@@

>>>>

The Kolmogorov-Smirnov (KS) statistic is widely used to test if a sample is
from a given distribution. This study demonstrates how block bootstrap can be 
used to approximate the KS statistic in situations where parameters are 
not specified and the data are serially dependent. We demonstrate the 
theoretical justication for this method, then through simulation we verify that
it holds it size under the null distribution and is powerful when the null 
hypothesis is false. Finally, we demonstrate applications of the method to
precipitation data from JFK airport and Microsoft stock returns.

@@

\paragraph{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ from a certain distribution $F(\theta)$,
and use the block bootstrap KS test to test if $X_t$ is generated from 
distribution $F$ with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformally distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
The results of the method depend on the sample size, so we must find the minimum
sample size required for the block bootstrap KS test to work.
We can implement the procedure described in methods with the true distribution
$F ~ N$, hypothesize distribution $F_0 ~ N$, and true mean $\mu = 8$ and 
standard deviation $\sigma = \sqrt{8}$.
We replicate this procedure 1000 times to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.
Then, we repeat the same procedure with $F ~ \Gamma$, and true shape $k = 8$ and
scale $\theta = 1$.

We constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, this 
indicates that the time series follows the null distribution. 
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$. 
%We can also
%compute the rate of rejection $\#P < \alpha$ for different values of
%$\alpha$, as well as a confidence interval
%for this proportion.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/normal}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $N(8,8)$ data generating process is normal.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/zoom_normal}
  \caption{A version of the above plot zoomed in to probabilities between 0 and
  0.1}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/gamma}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $\gamma(8,1)$ data generating process is gamma distributed.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/zoom_gamma}
  \caption{A version of the above plot zoomed in to probabilities between 0 and
  0.1}
  \label{fig:mu}
\end{figure}

\paragraph{Power}
We must also evaluate if this method works when $H_0$ is false. We can use
the block bootstrap KS test to test if $X_t$ is generated from some other 
distribution $G$ with some unknown $\theta$. In this scenario, if the test is 
indeed powerful,
we would ideally want $\beta$, or the probability of rejection to be 1, but we
expect it to be generally high. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the rate
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

If we can show that the test correctly fails to reject $H_0$ under the null
distribution, and correctly rejects $H_0$ for an alternate distribution, we can
state that the method works.

We can find the rate of rejection for  different values for the significance 
level $\alpha$. Additionally, we can construct a 95\% confidence interval for 
this rate. Below are rejection rates for $\phi = 0.2$ and $n = 800$, when
applied to the normal and gamma distributions using $\alpha =$ 0.01, 0.05, and 
0.10.

\input{tables/rr}

As shown in the above table, the rejection rates for these settings are very 
high (almost 1), indicating that the test is very powerful for these settings.

@@

\subsection{Precipitation in Mansfield Connecticut}
Monthly maximums of daily precipiation in inches time series data was retrieved 
from the John F. Kennedy (JFK) International Airport station from the 
National Oceanic and Atmospheric Administration.
%\mc{https://hdsc.nws.noaa.gov/pfds/pfds_map_cont.html?bkmrk=ct}

@@

We have shown that given a large enough sample of a time series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. It requires a larger sample size than 
basic bootstrap.
